Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-03-28 WSSSPE2_report/WSSSPE2_review.tex
Section 10: Add a question in note.
edits in response to edits and comments

\begin{table*}[th]
\begin{center}
\caption{Enablers of software sustainability, with 0 to 10 `*'s roughly
indicating the fraction of attendees who voted for an item as important.}
    \begin{scriptsize}
    \begin{tabular}{ | p{1.65cm} | p{10.0cm} |}
    \hline
  Importance & item \\ \hline \hline
********** & healthy and vibrant communities; vibrant community to champion software \\ \hline
********** & designing for growth and extension --- open development \\ \hline
******* & culture in community for reuse \\ \hline
**** & portability \\ \hline
**** & culture in dev community to support transition between developers \\ \hline
*** & interdisciplinary people : science + IT experience \\ \hline
** & planning for end of life \\ \hline
** & make smart choices about dependencies \\ \hline
* & thinking of software as product lines --- long term vs.~short term view \\ \hline
 & not all communities need new software \\ \hline
 & converting use into resources \\ \hline
    \end{tabular}
    \end{scriptsize}
    \label{tb:software_sustainability_enablers}
\end{center}   
\end{table*}

\begin{table*}[ht]
\begin{center}
\caption{Barriers to software sustainability, with 0 to 10 `*'s roughly
indicating the fraction of attendees who voted for an item as important.}
    \begin{scriptsize}
    \begin{tabular}{ | p{1.65cm} | p{10.0cm} |}
    \hline
  Importance & item \\ \hline \hline
******* & lack of incentives, including promotion and tenure process; promotion
and tenure process in academic is incompatible with sustainability \\ \hline
*****  & absent or poor documentation \\ \hline
***** & funding to ensure sustainability is difficult to obtain \\ \hline
**** & developers are not computer scientists; don't have software engineering
practices (in particular, those needed to scale-up projects to support and be
developed by a large sustainable community) \\ \hline
*** & overreliance on one or two people - bus test \choinote{What does ``bus test'' mean?}\\ \hline
** & rate of change of underlying technologies \\ \hline
** & lack of business models for sustainability \\ \hline
* & lack of training for how to build sustainability into the system \\ \hline
 & maintenance needed for software is not visible, appears to ``just happen'' \\ \hline
 & licensing issues \\ \hline
 & staff turnover - lack of continuity \\ \hline
    \end{tabular}
    \end{scriptsize}
    \label{tb:software_sustainability_barriers}
\end{center}   
\end{table*}

@@

To enable a form of standardized crediting for developers of research software,
the group proposed to work towards a taxonomy for software-based
contributorship. The taxonomy should be derived from, or extend, existing
taxonomies for research impact and contributorship such as defined by CASRAI (in
particular based on the Wellcome-Harvard contributorship taxonomy\katznote{Dan
to add cite or url}), VIVO, or ISNI. An interesting measure of impact raised by
the group was the {\em betweenness centrality}, an indicator of a person's
centrality (and hence, importance) in a scientific collaboration. It is expected
that developers of research software often play such a central role.

@@

The WSSSPE2 workshop continues our experiment from WSSSPE1 in how we can
collaboratively build a workshop agenda. The differences in WSSSPE2 from WSSSPE1
are in using an existing service (EasyChair) to handle submissions and reviews,
rather than an ad hoc process, and using an existing service (well-sorted) to
allow collaborative grouping of papers into themes by all authors, reviewers,
and the community, rather than this being done in an ad hoc manner by the
organizers alone.

>>>>

\begin{table*}[th]
\begin{center}
\caption{Enablers of software sustainability, with 0 to 10 `*'s roughly
indicating the fraction of attendees who voted for an item as important.}
    \begin{scriptsize}
    \begin{tabular}{ | p{1.65cm} | p{10.0cm} |}
    \hline
  Importance & item \\ \hline \hline
********** & healthy and vibrant communities; vibrant community to champion software \\ \hline
********** & designing for growth and extension -- open development \\ \hline
******* & culture in community for reuse \\ \hline
**** & portability \\ \hline
**** & culture in dev community to support transition between developers \\ \hline
*** & interdisciplinary people : science + IT experience \\ \hline
** & planning for end of life \\ \hline
** & make smart choices about dependencies \\ \hline
* & thinking of software as product lines -- long term vs.~short term view \\ \hline
 & not all communities need new software \\ \hline
 & converting use into resources \\ \hline
    \end{tabular}
    \end{scriptsize}
    \label{tb:software_sustainability_enablers}
\end{center}   
\end{table*}

\begin{table*}[ht]
\begin{center}
\caption{Barriers to software sustainability, with 0 to 10 `*'s roughly
indicating the fraction of attendees who voted for an item as important.\katznote{note that there's a footnote within this table -- Where has LaTeX put it?  How do we get it to show up near here?}}
    \begin{scriptsize}
    \begin{tabular}{ | p{1.65cm} | p{10.0cm} |}
    \hline
  Importance & item \\ \hline \hline
******* & lack of incentives, including promotion and tenure process; promotion
and tenure process in academic is incompatible with sustainability \\ \hline
*****  & absent or poor documentation \\ \hline
***** & funding to ensure sustainability is difficult to obtain \\ \hline
**** & developers are not computer scientists; don't have software engineering
practices (in particular, those needed to scale-up projects to support and be
developed by a large sustainable community) \\ \hline
*** & overreliance on one or two people - the `bus test'\footnote{how few key people would the project need to lose to become non-viable?  (larger is better)} \\ \hline
** & rate of change of underlying technologies \\ \hline
** & lack of business models for sustainability \\ \hline
* & lack of training for how to build sustainability into the system \\ \hline
 & maintenance needed for software is not visible, appears to ``just happen'' \\ \hline
 & licensing issues \\ \hline
 & staff turnover - lack of continuity \\ \hline
    \end{tabular}
    \end{scriptsize}
    \label{tb:software_sustainability_barriers}
\end{center}   
\end{table*}

@@

To enable a form of standardized crediting for developers of research software,
the group proposed to work towards a taxonomy for software-based
contributorship. The taxonomy should be derived from, or extend, existing
taxonomies for research impact and contributorship such as defined by CASRAI (in
particular based on the Wellcome-Harvard contributorship taxonomy\footnote{Project CRediT, \url{http://credit.casrai.org/}}, VIVO, or ISNI. An interesting measure of impact raised by
the group was the {\em betweenness centrality}, an indicator of a person's
centrality (and hence, importance) in a scientific collaboration. It is expected
that developers of research software often play such a central role.

@@

The WSSSPE2 workshop continues our experiment from WSSSPE1 in how we can
collaboratively build a workshop agenda. The differences in WSSSPE2 from WSSSPE1
are in using an existing service (EasyChair) to handle submissions and reviews,
rather than an ad hoc process, and using an existing service (well-sorted) to
allow collaborative grouping of papers into themes by all authors, reviewers,
and the community, rather than this being done in an ad hoc manner by the
organizers alone. \choinote{Are lightning talks a new feature in WSSSPE2?} \katznote{yes, along with a change to breakout groups that were intended to define actions, then potentially become longer lasting working groups to perform those actions}
